---
title: "SDS_HW2"
output:
  pdf_document: default
  html_document: default
date: "2024-01-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, results='hide'}
library(tidyverse)
library(conflicted)
library(dplyr)
conflicts_prefer(dplyr::filter)
```

```{r echo =FALSE, results='hide'}
profs_data <- read.csv("/Users/disharaj/Downloads/profs.csv")
```

```{r echo = FALSE}
library(ggplot2)
ggplot(profs_data, aes(x = eval)) + geom_histogram(binwidth = 0.1, fill = "seashell", color = "orange") + labs(title = "Distribution of Course Evaluation Scores", x = "Evaluation Scores", y = "Frequency")+ theme_minimal()
```

This histogram depicts the instructor’s average teaching evaluation score, on a scale of 1 to 5, for
courses in the sample. These data represent evaluations from 25,547 students and most major
academic departments. The data frame also includes information on characteristics of each course
and facts
about the professors such as seniority and demographics.
```{r echo = FALSE}
ggplot(profs_data, aes(x = minority, y = eval)) + 
  geom_boxplot(fill = "seashell", color = "orange") +
  labs(title = "Distribution of Course Evaluation Scores by Native Speaker Status", 
       x = "Native Speaker",
       y = "Evaluation Score") +
  theme_minimal()
```

This boxplot shows the distribution of course evaluation scores by whether or not the
professor is a native English speaker.
```{r echo = FALSE}
ggplot(profs_data, aes(x = eval)) + 
  geom_histogram(binwidth = 0.1, fill = "seashell", color = "orange") +
  labs("Distribution of Evaluation scores by gender", x = "Evaluation Scores", y = "Frequency") + facet_wrap(~gender) 
```

This is a faceted histogram with two rows to compare the distribution of course evaluation scores for
male and female instructors.
```{r echo = FALSE}
ggplot(profs_data, aes(x = eval, y = beauty)) +
  geom_point(color = "orange") +
  labs(title = "Relationship between physical attractiveness and evaluation scores", x = "Evaluation Scores", y = "Physical Attractiveness") +
  theme_minimal()
```

This is a scatterplot to visualize the extent to which there may be an association between the
professor’s physical attractiveness (x) and their course evaluations (y).
```{r echo = FALSE, results='hide'}
bikes_data <- read.csv('/Users/disharaj/Downloads/bikeshare.csv')
```

```{r echo=FALSE}
bikes_data1 <- bikes_data %>%
  group_by(hr) %>%
  summarise(avg_total = mean(total, na.rm = TRUE))
ggplot(bikes_data1, aes(x = hr, y = avg_total)) +
  geom_line(color = "orange", linewidth = 1) +
  labs(title = "Average hourly bike rentals accross all hours of the day", x = "Hours of day", y = "Average hourly bike rentals")+  theme_minimal()
```

This data set contains a two-year historical log (2011 and 2012) from the Capital Bikeshare system in
Washington D.C. These data have been aggregated on an hourly and daily basis and then merged
with weather and seasonal data. This line graph shows the average hourly bike rentals (total) across
all hours of the day (hr).
```{r echo = FALSE}
bikes_data1 <- bikes_data %>%
 group_by(hr, workingday) %>%
 summarise(avg_total = mean(total, na.rm = TRUE)) 
ggplot(bikes_data1, aes(x = hr, y = avg_total)) +
  geom_line(color = "orange", linewidth = 1) +
  labs(title = "Average hourly bike rentals accross all hours of the day, according to weather it's a working day", x = "Hours of day", y = "Average hourly bike rentals") +
   facet_wrap(~workingday)
```

This is a faceted line graph showing average bike rentals by hour of the day, faceted according to
whether it is a working day (workingday).
```{r echo = FALSE}
bikes_data1 <- bikes_data %>%
  filter(hr ==9) %>%
  group_by(weathersit, workingday) %>%
  summarise(avg_ridership = mean(total, na.rm = TRUE))
ggplot(bikes_data1, aes(x = weathersit, y = avg_ridership)) +
  geom_bar(stat = "identity", fill = "seashell", color = "orange") + 
  labs(title = "Average rideship during the 9 A.M hour by weather, according to working day", x = "Weather Situation", y = "Average rideship") + 
  facet_wrap(~workingday) 
```
```{r echo = FALSE, results='hide'}
capmetro_data <- read.csv('/Users/disharaj/Downloads/capmetro_UT.csv')
```

This bar graph is a faceted bar plot showing average ridership (y) during the 9 AM hour by weather
situation, faceted according to whether it is a working day or not.
```{r echo = FALSE}
capmetro_data = mutate(capmetro_data,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))
```

```{r echo = FALSE}
capmetro_data1 <- capmetro_data %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarise(avg_boardings = mean(boarding, na.rm = TRUE))
ggplot(capmetro_data1, aes(x = hour_of_day, y = avg_boardings, color = month)) +
  scale_color_manual(values = c("Sep" = "orange", "Oct" = "coral", "Nov" = "peru")) +
  geom_line(size = 1) +
  labs(title = "Average boarding by hour for each month, for each day of the week", x = "Hour of day", y = "Average Boardings") +
  facet_wrap(~day_of_week)
```

These graphs track ridership on buses in the UT area. Ridership is measured by an optical scanner
that counts how many people embark and alight the bus at each stop. Each row in the data set
corresponds to a 15-minute period between the hours of 6 AM and 10 PM, each and every day, from
September through November 2018.
This faceted line graph plots average boardings by hour of the day, day of week, and month. Also
these graphs are faceted by day of week. Each includes three lines of average boardings (y) by hour
of the day (x), one line for each month and distinguished by color.
1. Yes, the average boardings varies from day to day. It is observed that, typically the boardings
are highest on Mondays through Wednesdays. It slows down from Thursdays to Fridays.
During weekends, it is very low, especially on Sunday, it is substantially low. Also observed
that during November, the average boardings are lower compared to other 2 months.
2. The boardings on Mondays during September is comparatively lower, which could be because of students' starting the semester in September.
3. Average boardings are lower on Wednesdays through Fridays during Novembers, which could be because it
is the exam season and more students would like to stay back home.
```{r echo =FALSE}
ggplot(capmetro_data, aes(x = temperature, y = boarding, color = weekend)) +
  scale_color_manual(values = c("weekend" = "orange", "weekday" = "peru")) +
  geom_point() +
  labs(title = "Relationship between boardings an temperature for different days of the week, by hour of day", x = "Temperature", y = "Boardings") +
 facet_wrap(~hour_of_day)
```

This is a faceted scatter plot showing boardings (y) vs. temperature (x), faceted by hour of the day,
and
with points colored in according to whether it is a weekday or weekend.
1. Yes, it is certainly observed that temperature does play a factor in determining the number
of UT students riding the bus. By keeping weekend status constant and hour of day constant,
it is observed that with increasing temperatures, more students ride the bus. We see more
footfall after temperatures go above 60 degrees. Similarly, the boardings drops as
temperatures soar above 90 degrees.
```{r echo=FALSE, results='hide'}
billboard_data <- read.csv('/Users/disharaj/Downloads/billboard.csv')
```

```{r echo = FALSE}
billboard_data1 <- billboard_data %>%
  filter(!is.na(weeks_on_chart)) %>%
  group_by(performer, song) %>%
  summarize(count = max(weeks_on_chart)) %>%
  select(performer, song, count) %>%
  arrange(desc(count))%>%
  head(10)
billboard_data1
```

The table above represents the top 10 most popular songs since 1958, which is determined on the basis of the number of weeks it spends on the Billboard Top 100. 
```{r echo = FALSE}
billboard_data1 <- billboard_data %>%
  filter(year != 1958 & year != 2021) %>%
  group_by(year) %>%
  summarize(unique_songs_count = length(unique(song)))
billboard_data1
```

```{r echo = FALSE}
ggplot(billboard_data1, aes(x = year, y = unique_songs_count)) +
  geom_line(color = "orange", linewidth = 1) +
  labs(title = "Musical Diversity Over the Years",
       x = "Year",
       y = "Number of Unique Songs on Billboard Top 100") +
  theme_minimal()
```

This line graph plots the measure of musical diversity over the years. The x axis shows the year, while
the y axis shows the number of unique songs appearing at any position on the Billboard Top 100 chart
in any week that year.
```{r echo = FALSE}
ten_week_hits <- billboard_data %>%
  group_by(performer, song) %>%
  filter(n() >= 10) %>%
  summarize(total_weeks = sum(weeks_on_chart)) %>%
  filter(total_weeks >= 10) %>%
  ungroup()
ten_week_hits_count <- ten_week_hits %>%
  group_by(performer) %>%
  summarize(ten_week_hits_count = n()) %>%
  filter(ten_week_hits_count >= 30)
ggplot(ten_week_hits_count, aes(x = reorder(performer, ten_week_hits_count), y = ten_week_hits_count)) +
  geom_bar(stat = "identity", fill = "seashell", color = "orange") +
  labs(title = "Number of Ten-Week Hits for Top Artists",
       x = "Performer",
       y = "Number of Ten-Week Hits") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This bar chart plots 19 artists in U.S. musical history since 1958 who have had at least 30 songs that
were “ten-week hits.”, showing how many ten-week hits each one had in their musical career.